A Test of the Frequency Independence Assumption of Power System Parameters used in Geomagnetically Induced Current Estimates

### 2019-06-26 paper-submitted-1.tex
Updates from Overleaf

\documentclass[draft,linenumbers]{agujournal2018}
\usepackage{apacite}
\usepackage{url} % should fix any errors with URLs in refs.
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage[colorlinks]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}

@@

\title{A Test of the Frequency Independence Assumption of Power System Coefficients used in Geomagnetically Induced Current Estimates}

@@

@@

\begin{keypoints}
\item Power system coefficients derived from a time series of GIC and geoelectric field measurements are frequency dependent.
\item A GIC model with frequency-dependent system coefficients provides significantly better estimates of GIC than a frequency independent model.
\item Data-derived power system coefficients may differ significantly with those computed using power system configuration information.
\end{keypoints}

\begin{abstract}
A common assumption used when estimating geomagnetically induced currents (GICs) in a power system given a time series of nearby direct measurements or indirect estimates of the horizontal geoelectric field components $E_x(t)$ and $E_y(t)$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency-independent power system coefficients.  A first test of this assumption uses GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away. A second model (Model 2) is obtained using a frequency domain generalization of Model 1: $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$. The coefficients $a(\omega)$ and $b(\omega)$ are computed from data and are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model~1. Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system coefficients may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters. It is also shown that further improvements over Model~2 can be made using a model with the geomagnetic field as an input.
\end{abstract}

@@

Geomagnetically induced currents (GICs) are electric currents in conducting systems that are due to electric fields near the Earth's surface. Time-varying electric and magnetic fields at Earth's surface are driven by time variations in ionospheric currents on time scales of minutes to hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}. Of particular interest are currents induced in electric power systems because they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012,Gaunt2014}. Accurate estimation of GIC magnitudes is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.

A common assumption made in estimates of GIC in an electric power system using either a measured or estimated time series of the horizontal geoelectric field at Earth's surface, $\mathbf{E}(t)$, is that the power system is resistive or quasi-DC in the sense that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}. In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression when contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.,][]{Boteler2014a,Boteler2014b,Boteler2017}. This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015}.

In this work, we use a unique dataset in which time series measurements of $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GIC was being measured in an electric power system.  We first estimate the system coefficients $a_o$ and $b_o$ from data using conventional methods, in which they are modeled as frequency independent. Then, we use a model in which the system coefficients are allowed to be frequency dependent, compute the frequency dependence, and compare this model's ability to represent GIC with the traditional frequency-independent model.

@@

The only modification made to the magnetic field measurements was the replacement of 36 data points on 2006-08-06 with linearly interpolated values. The 36 data points for $B_x$ and $B_y$ have the same value and are much larger than the surrounding values.

@@

In Model~2, the system coefficients $a(\omega)$ and $b(\omega)$ are frequency dependent according to

\begin{linenomath*}
\begin{equation}
G_E(\omega) = a(\omega)E_x(\omega) + b(\omega)E_y(\omega)
\label{model2}
\end{equation}
\end{linenomath*}
\noindent
and are estimated using the Fourier transforms of  $GIC(t)$ and $\mathbf{E}(t)$; Frequency-dependent variables are indicated by explicitly listing their dependence on $\omega$. The method of parameter estimation and the generation of time-domain GIC estimates for Models~2-4 are given at the end of this section.

Motivated by the results described in the following section in which Model~2 provides a significantly better representation of GIC than Model~1, we have also considered two additional modeling approaches that can be used to specify GIC. In these two models, a direct estimate of the geoelectric field is not needed; instead, the driver is the geomagnetic field, which is much more frequently measured, and as a result, these models may be of practical interest.

In Model~3, the coefficients $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field, $\mathbf{B}$, and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ and

@@

\noindent
where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. 
That is, instead of using the measured electric field directly as in Model~2, a magnetotelluric transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.

@@

\noindent
In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce estimates of GIC that are similar to that obtained from Model~3 because they both use the same input variable.

Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ in a band of frequencies around $f_e$ from the DFT (discrete Fourier transform) of 24 hours of 1-s cadence data. Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used. The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation window has a lower bound of $1/86400$~Hz and highest evaluation window has an upper bound of $0.5$~Hz.

@@

Transfer function coefficients were computed for each of the segments. The average transfer function magnitude and phase at a given evaluation frequency was obtained by averaging the segment values at that evaluation frequency.

To interpolate the frequency-domain transfer function coefficients estimated at the evaluation frequencies, their real and imaginary components are linearly interpolated individually onto the original DFT frequency grid (with a spacing of $1/86400$~Hz), and then the inverse DFT is computed. The interpolated transfer function coefficients were set to zero at frequencies higher than $0.25$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.

GIC time series were calculated by multiplication of the interpolated transfer function coefficients with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(f)=a^I(f)E_x(f)+b^I(f)E_y(f)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(f)$ was calculated to give the estimate of $G_E(t)$ on a 1~s time grid).

Several modifications to this basic procedure were considered. Robust regression \citep{Egbert1986} with a Huber weighting constant of $1.345$ and an upper limit of 50 iterations and bi-square weighting with a constant of $4.685$ (both constants provide 95\% efficiency for normally distributed errors), both with no hard-cutoff step; time domain pre-whitening and tapering; and weighting the frequencies around each evaluation frequency using Parzen window weights. When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method. Transfer function coefficients found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods higher than $\sim30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)

@@

Out-of-sample estimation metrics for each model on day $D$ were determined by computing the average model coefficients using the remaining 34 days of data and then using these averaged coefficients to generate GIC time series and associated metrics on day $D$. This process was repeated to create a set of 35 out-of-sample prediction metrics for each model. In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the estimate at the start of the interval for the amount of time that the causal part of the impulse response (found by Fourier inversion of the transfer function) is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.

The performance of each model in estimating GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio. The PE and MSE ratios are shown in Table~\ref{results}.

The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that is predicted by the model.

The average PE and MSE ratios were computed by averaging the out-of-sample PE and MSE ratios. The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the out-of-sample prediction metrics. (The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.)

The average signal-to-noise ratio at a given frequency is the average of the out-of-sample signal-to-noise ratios at that frequency. The spectra for the signal and prediction error used in the signal-to-noise ratio calculation were obtained by averaging the raw spectra in the same frequency windows used for the regression.

\begin{table}
\caption{Out-of-sample model performance metrics for models $m=1-4$. MSE$_1$/MSE$_m$ is the ratio of the MSE for Model~1 to that of Model $m$. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_1$/MSE$_m$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
\hline
\end{tabular}
\label{results}
\end{table}

Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~\ref{results}, i.e., Model~1 has the lowest PE and Model~4 has the highest. Note that the large errors visible at the start and end of the time intervals are due to the transients discussed above and the first and last 10~minutes of each interval were omitted from metrics calculations.

The primary conclusions of this work follow from the results shown in Table~~\ref{results}: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.

@@

Figure~\ref{SN} shows the average signal-to-noise (i.e., signal to prediction error) ratios as a function of period for the four models. At each evaluation frequency, the SN ratio is the ratio of the average of the out-of-sample SN ratios of the segments and the error bars represent a 95\% CIs based on 1000 bootstrap samples of the segment SN ratios. 
Consistent with the results in Table~\ref{results}, the SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies. However, ordering of the SN ratio for Models~2-4 is dependent on period.

Below $600$~s (10~min), Model~2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high-frequency fluctuations in the prediction error for Model~2 are the smallest among the models. For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a higher SN ratio than Model~3, which uses an indirect estimate of the electric field. Model~4 has a slightly higher SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the data-estimated coefficients in $\mathbf{Z}$.

@@

Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~\ref{results}, with Model~1 having the lowest and Model~4 having the highest SN ratio.

The average coefficients and 95\% CI for Model 1 are $a_o = 73$ $[62,83]$ A/(V/km) and $b_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the coefficients for the segments is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).

The average coefficients for Models~2 and~4 are shown in Figure~\ref{Z}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.

Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition). For this system, GIC is more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).

@@

Figure~\ref{Phi} shows the transfer function phases (that for Model~1 is 0$^o$ for positive coefficients and 180$^o$ for negative coefficients by definition). The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.

>>>>

\documentclass[draft,linenumbers]{agujournal2018}
\usepackage{apacite}
\usepackage{url} %this package should fix any errors with URLs in refs.
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage[colorlinks]{hyperref}
\usepackage[colorinlistoftodos]{todonotes}

%\usepackage{framebox}

%%%%%%%
%\usepackage{trackchanges}
% uncomment the line above to use the TrackChanges package to mark revisions if needed.
% The trackchanges package adds five new LaTeX commands:
%
%  \note[editor]{The note}
%  \annote[editor]{Text to annotate}{The note}
%  \add[editor]{Text to add}
%  \remove[editor]{Text to remove}
%  \change[editor]{Text to remove}{Text to add}
%
% complete documentation is here: http://trackchanges.sourceforge.net/

@@

% Not working
% \listoftodos

\title{A Test of the Frequency Independence Assumption of Power System Parameters used in Geomagnetically Induced Current Estimates}

@@

%Weigel: http://orcid.org/0000-0002-9521-5228
%Cilliers: http://orcid.org/0000-0003-3175-5134

@@

\begin{keypoints}
\item Power system parameters derived from a time series of GIC and geoelectric field measurements are frequency dependent.
\item A GIC prediction model with frequency-dependent system parameters provides significantly better predictions than a frequency independent model.
\item A model derived using GIC and geomagnetic field data provides better predictions than a GIC and geoelectric field-based model.
\end{keypoints}

\begin{abstract}
A common assumption used when estimating geomagnetically induced currents (GICs) in a power system given a time series of nearby direct measurements or indirect estimates of the horizontal geoelectric field components $E_x(t)$ and $E_y(t)$ on Earth's surface is that the system is resistive. That is, the approximation $GIC(t)$  $=$ $a_oE_x(t) + b_oE_y(t)$ (Model 1) is used, where $a_o$ and $b_o$ are frequency-independent network parameters.  A first test of this assumption is made using GIC measurements made in a 187 kV transformer connected to a $\sim$100~km power line in Memanbetsu, Japan and geoelectric field measurements made at the Memanbetsu Magnetic Observatory $\sim$9~km away. A second model (Model 2) is obtained using a frequency domain generalization of Model 1: $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$. The network parameters $a(\omega)$ and $b(\omega)$ are computed from data and are shown to be frequency dependent and this model provides significantly better estimates of the measured GIC than Model 1. A third model, which uses the electric field estimated by convolving a magnetotelluric transfer function with the geomagnetic field, provides significantly better estimates over Model~2. Based on results using a simulated geoelectric field, it is suggested that the measurement-derived frequency dependence of the system parameters may be explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.

%Network parameters $a(\omega)$ and $b(\omega)$ computed from data in Model 2, $GIC(\omega)$ $=$ $a(\omega)E_x(\omega) + b(\omega)E_y(\omega)$, are shown to be frequency dependent and this model provides a significantly better estimates of the measured GIC over Model~1.

%It is suggested that the measurement-derived frequency dependence of the system parameters is explained by spatial variations in the spectrum of the geoelectric field along the line in which GIC is measured rather than by an actual frequency dependence of the system parameters.

\end{abstract}

@@

Geomagnetically induced currents (GICs) are electric currents in conducting systems that are due to electric fields near the Earth's surface. Time-varying electric and magnetic fields at Earth's surface are driven by time variations in ionospheric currents on time scales on the order of hours \citep{Ohtani2000} and the movement of Earth's surface relative to slow-varying current systems in Earth's ionosphere that are near stationary relative to the Sun \citep{Stening2013}. Of particular interest are currents induced in electric power systems because they can lead to system degradation, disruption, and failure \citep{Albertson1993,NERC2012,Gaunt2014}. Accurate estimation of GIC magnitudes is important for power system design, retrospective analysis, and mitigation of the impacts of space weather on power systems \citep{Molinski2002,Thomson2010,NERC2012,Gaunt2014}.

A common assumption made in estimates of GIC in an electric power system using either a measured or estimated time series of the horizontal geoelectric field at Earth's surface, $\mathbf{E}(t)$, is that the power system is resistive or quasi-DC in the sense that the relationship $GIC(t) = a_oE_x(t) + b_oE_y(t)$ holds \citep{Albertson1981,Lehtinen1985}. In this case, estimates of the coefficients $a_o$ and $b_o$ can be easily made from the data using linear regression when contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$ are available, or by using information about the connectivity of the power lines and values of the conductor and transformer resistances using DC circuit methods \citep[e.g.,][]{Boteler2014a,Boteler2014b,Boteler2017}. This quasi-DC assumption has been used or stated in many GIC-related works, for example, \citet{Pulkkinen2007,Wik2008,Pulkkinen2010,Ngwira2011,Horton2012,Viljanen2012,Overbye2012,Marshall2013,Liu2014,Zheng2014,Watari2015,Bonner2017}.

%\cite{Oyedokun2013a} developed a method to account for the finite inductive response time of transformers by taking the GIC predicted based on the quasi-DC assumption and then applying a piecewise-linear L/R filter derived from bench-scale transformer measurements, with a time constant dependent on load, to obtain an alternative estimate of GIC.

%\todo[inline]{What can I conclude from \cite{Oyedokun2013a}? Mostly the thesis documents visual differences in time series traces with and without accounting for transformers.}

In this work, we use a unique dataset in which time series measurements of $\mathbf{E}$ and the surface horizontal geomagnetic field, $\mathbf{B}$, were made near a site where GIC was being measured in an electric power system.  We first estimate the system parameters $a_o$ and $b_o$ from data using conventional methods, in which they are modeled as frequency independent. Then, we use a model in which the system parameters are allowed to be frequency dependent, compute the frequency dependence, and compare this model's ability to represent GIC with the traditional frequency-independent model.

@@

The only modification made to the magnetic field measurements was the replacement of 36 data points on 2006-08-06 with linearly interpolated values. The 36 values for $B_x$ and $B_y$ were the same and much larger than the surrounding values.

@@

In Model 2, the system parameters $a(\omega)$ and $b(\omega)$ are frequency dependent according to

\begin{linenomath*}
\begin{equation}
G_E(\omega) = a(\omega)E_x(\omega) + b(\omega)E_y(\omega)
\label{model2}
\end{equation}
\end{linenomath*}
\noindent
and are computed using contemporaneous measurements of $GIC(t)$ and $\mathbf{E}(t)$. The method of parameter estimation and generating predictions for Models 2-4 are given at the end of this section and fourier transformed variables are indicated by explicitly listing their dependence on $\omega$.

In Model~3, the parameters $a(\omega)$ and $b(\omega)$ from Model~2 are used, but with an electric field $\mathbf{E}'$ computed using the measured magnetic field, $\mathbf{B}$, and a transfer function $\boldsymbol{Z}$: $\mathbf{E}'(\omega) \equiv \boldsymbol{Z}(\omega)\mathbf{B}(\omega)$ and

@@

\noindent
where, explicitly, the electric field components are $E'_x$ = $Z_{xx}B_x + Z_{xy}B_y$ and $E'_y = Z_{yx}B_x + Z_{yy}B_y$. 
That is, instead of using the measured electric field directly as in Model 2, a magnetotelluric transfer function that relates $\mathbf{E}$ to $\mathbf{B}$ is used to provide an estimate of the electric field based on the magnetic field.

%Substitution gives

%\begin{linenomath*}
%\begin{equation*}
%G_{E'}(\omega) = %\big[a(\omega)Z_{xx}(\omega) + b(\omega)Z_{yx}(\omega) \big] B_x(\omega) + \big[ a(\omega)Z_{xy}(\omega) + b(\omega)Z_{yy}(\omega) \big] B_y(\omega)\,.
%\end{equation*}
%\end{linenomath*}

@@

\noindent
In this model, the transfer function components $z_x$ and $z_y$ are determined directly from measurements of $GIC(t)$ and $\mathbf{B}(t)$ and no electric field measurements are used; this model is expected to produce predictions of GIC that are similar to that obtained from Model 3 because they both use the same input variable.

Estimates of $a(\omega)$ and $b(\omega)$ are made at evaluation frequencies, $\omega_e \equiv 2\pi f_e$, by performing a ordinary least squares regression (OLS) using Equation~\ref{model2} and a set $G_E(\omega)$ and $\mathbf{E}(\omega)$ values near $f_e$ from the DFT (discrete fourier transform) of 24 hours of 1-s cadence data. Following the suggestion of \cite{Simpson2005}, the 28 values of $f_e$ start at $0.25$ Hz and decrease by a factor of $\sqrt{2}$ and end at $1/43200$~Hz; if a $f_e$ value computed in this way is not an integer multiple of $1/86400$, the nearest smaller integer multiple is used. The regression for each $f_e$ uses all DFT frequencies with values in a window with a width of twice the separation from the next lower value of $f_e$ and the given $f_e$ with the exceptions that the lowest evaluation window has a lower bound of $1/86400$~Hz and highest evaluation window has an upper bound of $0.5$~Hz.

@@

Transfer function parameters were computed for each of the segments. The average transfer function magnitude and phase at a given evaluation frequency was obtained by averaging the segment values at that evaluation frequency.

To interpolate the frequency-domain transfer function parameters estimated at the evaluation frequencies, their real and imaginary components are linearly interpolated individually onto the original DFT frequency grid (with a spacing of $1/86400$~Hz), and then the inverse DFT is computed. The interpolated transfer function parameters were set to zero at frequencies higher than $0.25$~Hz and the transfer function estimate at $f=0$ was set to zero before interpolation.

Prediction time series were calculated by convolution of the interpolated transfer function parameters with the DFT of the 24-hr long and 1~s cadence model inputs and computing the inverse DFT (e.g, for Model~2, $G_E(f)=a^I(f)E_x(f)+b^I(f)E_y(f)$ was calculated, where the superscript $I$ indicates the interpolated value, and then the inverse DFT of $G_E(f)$ was calculated to give the prediction $G_E(t)$ on a 1~s time grid).

Several modifications to this basic procedure were considered. Robust regression \citep{Egbert1986} with a Huber weighting constant of $1.345$ and an upper limit of 50 iterations and bi-square weighting with a constant of $4.685$ (both constants provide 95\% efficiency for normally distributed errors), both with no hard-cutoff step; time domain pre-whitening and tapering; and weighting the frequencies around each evaluation frequency using Parzen window weights. When using these additional steps individually, we find that the conclusions made regarding the differences each model's prediction ability do not change because the differences in each model's prediction performance are within the calculated uncertainty of the used method. Transfer function parameter values found using these modifications are generally within the error bars of those for the used procedure when the signal to noise ratio is above $\sim$3 (corresponding to periods higher than $\sim30$~s; At periods below $\sim$30~s, the electric field measurements are significantly influenced by the instrument response characteristics~\citep{Fujii2015}, aliasing, and the computed values are also expected to be significantly biased due to low signal to noise ratio.)

@@

Out-of-sample prediction metrics for each model on day $D$ were determined by computing the average model parameters using the remaining 34 days of data and then using these averaged parameters to generate a prediction time series and associated metrics on day $D$. This process was repeated to create a set of 35 out-of-sample prediction metrics for each model. In calculating the metrics, the first and last 10 minutes of the day were excluded to reduce the influence of transients that exist in the prediction at the start of the interval for the amount of time that the causal part of the impulse response is non-zero and at the end for the amount of time that the acausal part of the impulse response is non-zero.

The performance of each model in predicting GIC was assessed using three metrics: the prediction efficiency, PE, the mean-squared-error (MSE) ratio, and the signal-to-noise ratio.

The prediction efficiency, PE, is the ``Case I'' skill score described by \cite{Murphy1988} and is defined as PE = 1 - variance(Prediction Error)/variance(Predicted); it is a measure of the skill of the model relative to the variance in the data that are predicted and it represents the fraction of the variance in the data that predicted by the model.

The average PE and MSE ratios were computed by averaging the out-of-sample PE and MSE ratios; the average signal-to-noise ratio at a given frequency is the average of the out-of-sample signal-to-noise ratios at that frequency. The 95\% confidence interval, CI, on the PE and MSE ratio was calculated using 1000 bootstrap samples of the out-of-sample prediction metrics. (The typical difference between the limits when the confidence intervals were computed assuming the values were Gaussian-distributed was $\sim$2\%.) The spectra for the signal and prediction error for each of the segments at a given evaluation frequency was obtained by averaging the raw spectra using the same frequency windows used for the regression.

\begin{table}
\caption{Out-of-sample model performance metrics for models $m=1-4$. MSE$_m$/MSE$_1$ is the ratio of the MSE for model $m$ to that of Model 1. The PE and MSE ratios are the average of 35 values and the confidence interval, CI, was determined from 1000 bootstrap samples.}
\centering
\begin{tabular}{l c c c c}
\hline
$m$\hspace{1em} Model & PE & 95\% CI & MSE$_m$/MSE$_1$ & 95\% CI\\
\hline
1\hspace{1em} $G_o(t) = a_oE_x(t) + b_oE_y(t)$ & 0.35 & [0.25, 0.45] & 1.0 & \\
2\hspace{1em} $G_E(t) = a(\omega)E_x(t) + b(\omega)E_y(t)$ & 0.60 & [0.54, 0.65] & 1.8 & [1.6, 2.0]\\
3\hspace{1em} $G_{E'}(t) = a(\omega)E'_x(t) + b(\omega)E'_y(t)$ & 0.78 & [0.76, 0.80] & 3.1 & [2.8, 3.4]\\
4\hspace{1em} $G_{B}(t) = Z_xB_x(t) + Z_yB_y(t)$ & 0.83 & [0.80, 0.85] & 4.6 & [3.8, 5.3]\\
\hline
\end{tabular}
\label{results}
\end{table}

Figure~\ref{predictions} shows the predictions for each model in a 1-day interval selected because the value of the prediction efficiency for each model was in the same order as that shown in Table~1, i.e., Model 1 has the lowest PE and Model 4 has the highest. Note that the large errors visible at the start and end of the time intervals are due to the transients discussed above and the first and last 10~minutes of each interval were omitted from metrics calculations.

The primary conclusions of this work follow from the results shown in Table~1: the frequency dependent model, Model~2, is significantly better at predicting GIC than the frequency independent model, Model~1; Model~3, which uses an estimate of the electric field based on magnetic field measurements, provides significantly better estimates of GIC than Model~2; Model~4, in which a transfer function that connects GIC to $\mathbf{B}$ was derived directly from the data, provides better estimates of GIC than Model~3.

@@

%Although inferior, the Model~1 does have a good prediction performance - the PE is 0.35 and the data/model correlation coefficient of 0.64. The top panel of  Figure~\ref{predictions} shows that its predicted GIC generally follows the overall trend of the measured GIC.

Figure~\ref{SN} shows the average signal-to-noise (i.e., signal to prediction error) ratios as a function of period for the four models. At each evaluation frequency, the SN ratio is the ratio of the average of the out-of-sample SN ratios of the segments and the error bars represent a 95\% CIs based on 1000 bootstrap samples of the segment SN ratios.

Consistent with the results in Table~1, the SN ratio for Model~1 is lower than that for Models~2-4 at all frequencies. However, ordering of the SN ratio for Models~2-4 is dependent on period.

Below $600$~s (10~min), Model~2 has the largest SN ratio. This result is somewhat visible in the sample interval shown in Figure~\ref{predictions} where the high-frequency fluctuations in the prediction error for Model~2 are the smallest among the models. For these periods, the predictive ability of each model is in the order that may be expected based on the data used to derive the model: Model~2, which uses direct measurements of the electric field, has a higher SN ratio than Model~3, which uses an indirect estimate of the electric field. Model~4 has a slightly higher SN than Model~3, and this is expected because they both use the same input of the magnetic field, but Model~3 has an additional source of uncertainty from the parameters in $\mathbf{Z}$.

@@

Above $10,000$~s (2.8~hrs), the SN ratio ordering is the same as the ordering of the PE and MSE ratios shown in Table~1, with Model~1 having the lowest and Model~4 having the highest SN ratio.

The average parameters and 95\% CI for Model 1 are $a_o = 73$ $[62,83]$ A/(V/km) and $b_o = -72.0$ $[-98,-45]$ A/(V/km) and the distribution of the parameters for the segments is shown in Figure~\ref{histogram}. (Results are similar if 1-minute averages are used; \cite{Watari2015} found $a_o=38.1$ A/(V/km) and $b_o=-7.4$ A/(V/km) for the interval 2006-12-14--2006-12-15 JST time interval).

%However, for the data considered in this work, when a model $G(t) = $ $c_oB_x(t) + d_oB_y(t)$ is used, were $c_o$ and $d_o$ are constants, the in-sample average prediction efficiency is comparable to that of Model 1, but the out-of-sample prediction efficiency is negative.

The average parameters for Models~2 and~4 are shown in Figure~\ref{Z}--\ref{Phi}. The error bars at each time step or period is the standard error of the 35 values used to compute the average.

%Figure~\ref{H} shows that the timescale of variations of an impulse in $\mathbf{E}$ for Model~2 and $\mathbf{B}$ for Model~4 are similar and are different from zero at more than one time value (the impulse response for Model~3 is similar to that of Model~2 and is not shown). By definition, the impulse response for Model~1 is zero except at $t=0$. Note that the sign of the $E_y$ impulse response at $t=0$ for Model~1 is opposite that for Model~2 (at $t=0$, $b_o < 0$ while $h_b(t=0)>0$). (An explanation for the acausality of the impulse response is given in \cite{Egbert1992,Kelbert2017}).

Figure~\ref{Z} shows that the frequency domain transfer functions for Model~2 vary by an order of magnitude on timescales of 60~s to 12~h (that for Model~1 is constant by definition). For this system, GIC more sensitive to $E_y$ than $E_x$ due to $b(\omega) > a(\omega)$, but note that $E_y(t)$ has less variability than $E_x(t)$ - the variance in $E_x(t)$ is $\sim$5x larger than $E_y(t)$ over the entire dataset (this is also visible in Figure~\ref{sample}).

@@

Figure~\ref{Phi} shows the transfer function phases (that for Model~1 is 0$^o$ for positive parameters and 180$^o$ for negative parameters by definition). The phases for $z_x(\omega)$ and $b(\omega)$ are offset by 120$^o$ while the phases for $a(\omega)$ and $b(\omega)$ are significantly different.

Figure~\ref{Z_MT} and~\ref{Phi_MT} show the parameters for $\mathbf{Z}$ in Model~3. Note that electric and magnetic field data from Memambetsu, Japan were also used by \cite{Fujii2015} to compute $\mathbf{Z}$. When using the \cite{Fujii2015} transfer function for Model~3, the PE was significantly lower (0.58 vs. 0.78). (The \cite{Fujii2015} transfer function was obtained from the IRIS DMC EMTF database [\cite{Kelbert2011}]). Although not critical to the issues addressed in this work, a preliminary investigation suggests that the cause is due to several factors. We have found that when using continuous one-month intervals prior to 2006 to compute $\mathbf{Z}$, our calculated transfer function matches the \cite{Fujii2015} transfer function to within the error bars except below periods of 100~s, which is likely explained by the issues with the electric field response in the 1-second-cadence measurements discussed in \cite{Fujii2015}. After 2006, our calculated transfer functions differ significantly from that in \cite{Fujii2015} - they are shifted downward by approximately 10-50\%. In addition, we have found that the estimated transfer function is sensitive to whether or not segment averaging was used. Due to the large gaps between and the varying length of continuous data segments used in this work, transfer function parameters were computed using one-day segments of data and then averaged. We have found that if a continuous set of measurements over approximately 21 days is used to compute a single transfer function, the transfer function magnitudes are up to 25\% higher than if 21 transfer functions were computed and then averaged.

%As a result, the transfer function parameters are shown in Figure~\ref{Z_MT} and~\ref{phi_MT} should be considered a

%\cite{Pulkkinen2010} noted that one could estimate the ratio of $a_o/b_o$ using GIC and local geomagnetic field measurements even if $\mathbf{Z}(\omega)$ was unknown if one assumed the ground conductivity was 1-D so that $Z\equiv Z_{xy}=-Z_{yx}$ and $Z_{xx}=Z_{yy}=0$ so that Equation~\ref{model2} can be written
